The CrypTool Script: \\ Cryptography, Mathematics and More \\ \vspace{30pt} {\large Background reading\\%*[-8pt] for the free e-learning program CrypTool\\ (10th edition -- distributed with CrypTool version 1.4.30)} \vspace{60pt}

### 2009-11-27 script/de/chapters/what-is-sage.tex
Script[de+en]: Many small changes, e.g. Sage samples collected in files, use of iload; described how to deal with Sage notebook using interact; substituted href by url.

Dieses Skript enthält zahlreiche mit Sage erstellte Programmbeispiele. Sage ist
ein Open-Source Computer-Algebra-System (CAS), das für Lehre, Studium und Forschung
in Mathematik eingesetzt wird.
Sage kombiniert viele hochwertige Open-Source-Packete\footnote{%
Einen Eindruck von der Größe von Sage erhält man, wenn man es selbst compiliert:
Die heruntergeladenen Sourcen von Sage 4.1 brauchten zur Compilierung auf einem
durchschnittlichen Linux-PC rund 5 h (inklusive aller Bibliotheken).
Danach nahm es 1,8 GB Plattenplatz ein.
}
und liefert den Zugang zu deren Funktionalität über ein gemeinsames, auf der
Programmiersprache Python basierendes Interface\footnote{%
Es gibt auch ein relativ einfaches Interface für die Sprache C, genannt Cython,
mit der man eigene Funktionen in Sage stark beschleunigen kann.\\
Siehe \url{http://openwetware.org/wiki/Open_writing_projects/Sage_and_cython_a_brief_introduction}.
}.

@@

\noindent Es gibt inzwischen viele PDF- oder HTML-Dokumente über Sage:
Ein guter Startpunkt ist die \glqq Bibliothek\grqq von der Sage-Hauptseite,
das \glqq Dokumentationsprojekt\grqq und das \glqq Lehrmaterial\grqq\footnote{%
- \glqq Bibliothek\grqq: {\centering \url{http://www.sagemath.org/library/index.html}},\\
- \glqq Dokumentationsprojekt\grqq:
  {\centering \url{http://wiki.sagemath.org/DocumentationProject}},\\
- \glqq Lehrmaterial\grqq: {\centering \url{http://wiki.sagemath.org/Teaching_with_SAGE}}.
}.

@@

% ---------------------------------------------------------------------------
\section*{Sage-Benutzerschnittstellen}
Sage ist kostenlos und kann von folgender Webseite herunter geladen werden:
\begin{center}
  \url{http://www.sagemath.org} \\
\end{center}
Standardmäßig nutzt man die Sage-Kommandozeile als Interface, wie im folgenden
Bild~\ref{fig:sage_cmd_interfaces} zu sehen ist.
Es gibt jedoch auch ein grafisches Benutzerinterface für diese Software
in Form des Sage-Notebooks (siehe Bild~\ref{fig:sage_gui_interfaces}).
Und schließlich kann man Sage auch online auf verschiedenen Servern nutzen, z.B.
\begin{center}
\url{http://www.sagenb.org} oder \\
\url{http://sage.mathematik.uni-siegen.de:8000}
\end{center}
ohne Sage lokal zu installieren.

@@

\footnotetext{%
Um das grafische Interface von Sage zu starten, muss man auf der Sage Kommandozeile
\verb#notebook()# eingeben. Danach startet der eingestellte Browser (Iceweasel,
Firefox, IE, ...) z.B. mit der URL \url{http://localhost:8000}.
}

@@

\noindent Danach kann man auf der Kommandzeile alles benutzen, was im
Sage-Skript definiert wurde\footnote{%
Anmerkungen:

@@

\noindent\hangindent=6pt\makebox[6pt][l]{-}Beim Laden des obigen
Sage-Skripts wird es von Sage zuerst geparst, und dann
in eine andere Datei namens \glqq primroots.py\grqq~kopiert. Sage ergänzt
dann alle notwendigen Variablen in \glqq primroots.py\grqq~und alle Import-Statements.
Somit wird das Sage-Skript genauso ausgeführt, als hätte man die Befehle einzeln
auf der Kommandozeile eingetippt. Ein Unterschied mag sein, dass Ausgaben explizit
mit \glqq print\grqq~erfolgen müssen.

@@

Normalerweise editiert man ein eigenes Sage-Skript wieder und möchte dann den
Inhalt des geänderten Skripts wieder in Sage laden. Dafür kann man den Befehl
\verb!attach()! nutzen. In unserem obigen Beispiel sieht das auf der Sage-Kommandozeile
folgendermaßen aus (man kann auch direkt nach dem \verb!load()! das \verb!attach()! aufrufen, und
nicht erst, wenn man das Skript ändert):

@@

\vspace{30pt}
\noindent Weitere Hinweise:
\begin{itemize}
  \item Abfrage der Version Ihrer Sage-Umgebung mit: \texttt{version()}
  \item Um sich schnell die Sage-Programmbeispiele in diesem Skript anzusehen, können Sie
  \begin{itemize}
    \item im Index nach \verb#Sage -> Programmbeispiele# schauen, oder 
    \item sich im Anhang das \glqq \nameref{sc:List-of-Sage-Code-Examples}\grqq~ansehen.
  \end{itemize}
\end{itemize}

>>>>

Dieses Skript enthält zahlreiche mit Sage erstellte Programmbeispiele. Sage ist
ein Open-Source Computer-Algebra-System (CAS), das für Lehre, Studium und Forschung
eingesetzt wird.
Sage kombiniert viele hochwertige Open-Source-Packete\footnote{%
Einen Eindruck von der Größe von Sage erhält man, wenn man es selbst compiliert:
Die heruntergeladenen Sourcen von Sage 4.1 brauchten zur Compilierung auf einem
durchschnittlichen Linux-PC rund 5 h (inklusive aller Bibliotheken).
Danach nahm es 1,8 GB Plattenplatz ein.
}
und liefert den Zugang zu deren Funktionalität über ein gemeinsames, auf der
Programmiersprache Python basierendes Interface\footnote{%
Es gibt auch ein relativ einfaches Interface für die Sprache C, genannt Cython,
mit der man eigene Funktionen in Sage stark beschleunigen kann.\\
Siehe \url{http://openwetware.org/wiki/Open_writing_projects/Sage_and_cython_a_brief_introduction}.
}.

@@

\noindent Es gibt inzwischen viele PDF- und HTML-Dokumente über Sage, so dass wir als
guten Startpunkt nur einige wenige nennen\footnote{%
- \glqq Bibliothek\grqq: {\centering \url{http://www.sagemath.org/library/index.html}},\\
- \glqq Dokumentationsprojekt\grqq:
  {\centering \url{http://wiki.sagemath.org/DocumentationProject}},\\
- \glqq Lehrmaterial\grqq: {\centering \url{http://wiki.sagemath.org/Teaching_with_SAGE}}.
}.

@@

% ---------------------------------------------------------------------------
\section*{Sage-Benutzerschnittstellen}
Sage ist kostenlos und kann von folgender Webseite herunter geladen werden:
\begin{center}
  \url{http://www.sagemath.org} \\
\end{center}
Standardmäßig nutzt man die Sage-Kommandozeile als Interface, wie im folgenden
Bild~\ref{fig:sage_cmd_interfaces} zu sehen ist.
Es gibt jedoch auch ein grafisches Benutzerinterface für diese Software
in Form des Sage-Notebooks (siehe Bild~\ref{fig:sage_gui_interfaces}).
Und schließlich kann man Sage-Notebooks\footnote{%
Weitere Details zu Sage-Notebooks finden Sie in
Kapitel~\ref{ec:Sage_Massierer} 
(\glqq \nameref{ec:Implementing-for-Education}\grqq
  $\Rightarrow$ \glqq \nameref{ec:Sage_Massierer}\grqq )
                      }
auch online auf verschiedenen Servern nutzen, ohne Sage lokal zu installieren, z.B.:
\begin{center}
\url{http://www.sagenb.org} oder \\
\url{http://sage.mathematik.uni-siegen.de:8000}
\end{center}
.

@@

\footnotetext{%
Um das grafische Sage-Interface lokal zu starten, muss man auf der Sage Kommandozeile
\verb#notebook()# eingeben. Danach startet der eingestellte Browser (Iceweasel,
Firefox, IE, ...) z.B. mit der URL \url{http://localhost:8000}.
}

@@

\noindent Danach kann man auf der Kommandozeile alle Variablen und Funktionen
nutzen, die im Sage-Skript definiert wurden\footnote{%
Anmerkungen:

@@

\noindent\hangindent=6pt\makebox[6pt][l]{-}Beim Laden des obigen
Sage-Skripts wird es von Sage zuerst geparst, und dann
in eine andere Datei namens \glqq primroots.py\grqq~kopiert. Sage ergänzt
dann alle notwendigen Variablen in \glqq primroots.py\grqq~und alle Import-Statements.
Somit wird das Sage-Skript genauso ausgeführt, als hätte man die Befehle einzeln
auf der Kommandozeile eingetippt. Ein bedeutender Unterschied ist,
dass alle Ausgaben ein \verb!print!~benötigen.

@@

Normalerweise editiert man ein eigenes Sage-Skript wieder und möchte dann den
Inhalt des geänderten Skripts wieder in Sage laden. Dafür kann man den Befehl
\verb!attach()! nutzen (man kann auch direkt nach dem \verb!load()! 
das \verb!attach()! aufrufen, und nicht erst, wenn man das Skript ändert;
man kann \verb!load()! sogar weglassen, da dies in \verb!attach()! enthalten ist):

@@

\vspace{20pt}
Falls man die Ausgabe einer attachten Datei so angezeigt haben möchte,
wie wenn man die Einzelbefehle direkt auf der Kommandozeile eingibt
(also nicht nur das, was per \verb!print! ausgegeben wird),
kann man den Befehl \verb!iload()! verwenden:
Jede Zeile wird dann einzeln geladen. Um die nächste Zeile zu laden,
muss man die \verb!Enter!-Taste drücken. Das muss man so lange wiederholen,
bis alle Zeilen des Sage-Skripts in die Sage-Session geladen sind.

\texttt{sage: iload primroots.sage}

% \vspace{30pt}
\newpage
\noindent Weitere Hinweise:
\begin{itemize}
  \item Abfrage der Version Ihrer Sage-Umgebung mit: \texttt{version()}
  \item Um sich schnell die Sage-Programmbeispiele in diesem Skript anzusehen, können Sie
    \begin{itemize}
      \item im Index nach \verb#Sage -> Programmbeispiele# schauen, oder 
      \item sich im Anhang das \glqq \nameref{sc:List-of-Sage-Code-Examples}\grqq~ansehen.
    \end{itemize}
  \item Der Quellcode der Sage-Beispiele in diesem Script wird in Form von
        Sage-Programmdateien mit dem CrypTool-Setup-Programm ausgeliefert.
        Nach der Installation finden Sie diese im Unterverzeichnis \verb#sage# innerhalb des
        CrypTool-Verzeichnisses:\\
        - SAGE-Samples-in-Chap01.sage \\
        - SAGE-Samples-in-Chap02.sage \\
        - SAGE-Samples-in-Chap03.sage \\
        - SAGE-Samples-in-Chap04.sage
\end{itemize}

### 2009-11-27 script/de/script-de.tex
fixed math mode in headings resulting in broken PDF bookmarks with \texorpdfstring; fixed float warnings by replacing placement [h] with [ht]; fixed warning about PDF version 1.5 by changing it to 1.4 in coverv1_*.pdf
enable verbose in footnotes using \VerbatimFootnotes; fixed some TODOs
CT script: Sage parts finished in (en); translation to German almost finished
fixed overfull hboxes
document used sw on page 2
Script[de+en]: hanging indent in footnotes; corrected links to old appendix parts
added contents overview based on shorttoc
Script[de+en]: Many small changes, e.g. Sage samples collect

\documentclass[a4paper,11pt,oneside]{book}
%\overfullrule=5pt % uncomment to mark overfull boxes
\usepackage[latin1]{inputenc} % Umlaute
% \usepackage[dvips]{graphics}
% \usepackage{graphicx}

@@

% \usepackage{flafter} Versuch von Hrn. Koy, zu erzwingen, dass 
%     Grafiken/Verschiebbare Objekte (floats) immer an der Stelle eingefügt
%     werden, wo sie aufgerufen werden. Tat nicht.

@@

% Formatierung der Fußnoten
\usepackage[flushmargin]{footmisc}
\usepackage{fullpage}
\usepackage[hyphens]{url} % options "hyphens" enables linebreaks in \url{}
% Literaturverzeichnis kapitelweise
\usepackage[labelstoglobalaux]{bibunits} % ohne labelstoglobalaux scheint man thebibilograpy und bibtex nicht mischen zu können
\usepackage{float}
\usepackage[colorlinks=true,linkcolor=black,citecolor=black,urlcolor=black,bookmarksnumbered=true,pdfpagelabels,plainpages=false,hyperfootnotes=false]{hyperref}
%To reduce the number of hyperlink related 'pdfTeX warning's we use options 
%pdfpagelabels, plainpages=false and, hyperfootnotes=false and place hyperref after float.
%Option debug helps finding problems.

\floatstyle{ruled}
\newfloat{sagecode}{!h}{loc}[chapter]
\floatname{sagecode}{Sage-Beispiel}
\newfloat{cryptoprocedure}{!h}{lop}[chapter]
\floatname{cryptoprocedure}{Krypto-Verfahren}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

@@

\vspace{450pt}
Quelle Coverfoto: \url{www.photocase.com}, Andre Günther

@@

\newpage
\pdfbookmark[0]{\contentsname}{Contents}
\tableofcontents

@@

\clearpage\phantomsection
\addcontentsline{toc}{chapter}{Verzeichnis der Sage-Programmbeispiele}
\listof{sagecode}{Verzeichnis der Sage-Programmbeispiele}

>>>>

\documentclass[a4paper,11pt,oneside]{book}
\overfullrule=5pt % uncomment to mark overfull boxes
\usepackage[latin1]{inputenc} % Umlaute
% \usepackage[dvips]{graphics}
% \usepackage{graphicx}

@@

@@

% Formatierung der Fußnoten
\usepackage[flushmargin]{footmisc}
\usepackage{fullpage}
\usepackage[hyphens]{url} % options "hyphens" enables linebreaks in \url{}
% Literaturverzeichnis kapitelweise
\usepackage[labelstoglobalaux]{bibunits} % ohne labelstoglobalaux scheint man thebibilograpy und bibtex nicht mischen zu können
\usepackage{float}
\usepackage{shorttoc}
\usepackage[colorlinks=true,linkcolor=black,citecolor=black,urlcolor=black,bookmarksnumbered=true,pdfpagelabels,plainpages=false,hyperfootnotes=false]{hyperref}
%To reduce the number of hyperlink related 'pdfTeX warning's we use options 
%pdfpagelabels, plainpages=false and, hyperfootnotes=false and place hyperref after float.
%Option debug helps finding problems.

\floatstyle{ruled}
\newfloat{sagecode}{!ht}{loc}[chapter]
\floatname{sagecode}{Sage-Beispiel}
\newfloat{cryptoprocedure}{!ht}{lop}[chapter]
\floatname{cryptoprocedure}{Krypto-Verfahren}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\VerbatimFootnotes

@@

\vspace{420pt}
Quelle Coverfoto: \url{www.photocase.com}, Andre Günther\\

Schriftsatz-Software: \LaTeX

Versionsverwaltungs-Software: Subversion

@@

\newpage
\pdfbookmark[0]{Kurzinhaltsverzeichnis}{ShortContents}
\shorttoc{Kurzinhaltsverzeichnis}{0}
\clearpage\phantomsection
\pdfbookmark[0]{\contentsname}{Contents}
\tableofcontents

@@

\clearpage\phantomsection
\addcontentsline{toc}{chapter}{Verzeichnis der Sage-Programmbeispiele}
\listof{sagecode}{Verzeichnis der Sage-Programmbeispiele}
\label{sc:List-of-Sage-Code-Examples}

\vspace{30pt}
\begin{itemize}
  \item Alle Beispiel wurden mit Sage v 4.2.1 (Release Date 2009-11-14) getestet.
  \item Der Quellcode der Sage-Beispiele in diesem Script wird in Form von
        Sage-Programmdateien mit dem CrypTool-Setup-Programm ausgeliefert.
        Nach der Installation finden Sie diese im Unterverzeichnis \verb#sage# innerhalb des
        CrypTool-Verzeichnisses:\\
        - SAGE-Samples-in-Chap01.sage \\
        - SAGE-Samples-in-Chap02.sage \\
        - SAGE-Samples-in-Chap03.sage \\
        - SAGE-Samples-in-Chap04.sage
\end{itemize}

### 2009-11-27 script/en/chapters/authors.tex
Script[de+en]: Added current research about twin primes (chap 3.8.3); made some small layout changes
Script[de+en]: Many small changes, e.g. Sage samples collected in files, use of iload; described how to deal with Sage notebook using interact; substituted href by url.

\item[Bernhard Esslinger,] \mbox{}\\
initiator of the CrypTool project, main author of this script, head IT
security at Deutsche Bank and professor for IT security at the University
of Siegen.
E-mail: besslinger@web.de, esslinger@fb5.uni-siegen.de.

@@

\item[Matthias B\"uger,] \mbox{}\\ 
contributor to the chapter ``Elliptic Curves'', research analyst at
Deutsche Bank.

@@

\item[Roger Oyono, ] \mbox{}\\
implementer of the CrypTool factorization dialog and
original author of chapter ``The Mathematical Ideas behind Modern
Cryptography''.

@@

\item[Johannes Buchmann,] \mbox{}\\
Co-author of the chapter ``Crypto2020 --- Perspectives for long-term cryptographic security''.
%Vice-president of the Technische Universit\"at Darmstadt, Germany and
%Professor at the Departments of Computer Science and of Mathematics.
Johannes Buchmann holds the Chair for Theoretical Computer Science (Cryptography
and Computer Algebra) at the department of Computer Science of the Technische
Universit\"at Darmstadt TUD).  He is also a Professor at the department of Mathematics, and vice-president of the university.

\item[Alexander May,] \mbox{}\\
Co-author of the chapter ``Crypto2020 --- Perspectives for long-term cryptographic security'', 
Assistant Professor at the department of Computer Science of
the Technische Universit\"at Darmstadt, Germany.

\item[Erik Dahmen,] \mbox{}\\
Co-author of the chapter ``Crypto2020 --- Perspectives for long-term cryptographic security'', 
%Researcher at the Department of Computer Science of the TUD.
Researcher at the Chair for Theoretical Computer Science (Cryptography and
Computer Algebra), department of Computer Science, Technische Universit\"at
Darmstadt, Germany.

\item[Ulrich Vollmer,] \mbox{}\\
Co-author of the chapter ``Crypto2020 --- Perspectives for long-term cryptographic security'', 
Researcher at the Chair for Theoretical Computer Science (Cryptography and
Computer Algebra), department of Computer Science, Technische Universit\"at
Darmstadt, Germany.

@@

\vskip +60 pt
\subsection{For Kids}

>>>>

\item[Bernhard Esslinger,] \mbox{}\\
initiator of the CrypTool project, main author of this script, head Information
Security Management at Deutsche Bank and professor for IT security and cryptography
at the University of Siegen.
E-mail: besslinger@web.de, esslinger@fb5.uni-siegen.de.

@@

\item[Matthias B\"uger,] \mbox{}\\ 
contributor to chapter~\ref{Chapter_EllipticCurves} (``\nameref{Chapter_EllipticCurves}''),
research analyst at Deutsche Bank.

@@

\item[Roger Oyono, ] \mbox{}\\
implementer of the CrypTool factorization dialog and original author
of chapter~\ref{Chapter_ModernCryptography} (``\nameref{Chapter_ModernCryptography}'').

@@

\item[Johannes Buchmann,] \mbox{}\\
Co-author of chapter~\ref{Chapter_Crypto2020} (``\nameref{Chapter_Crypto2020}'').
%Vice-president of the Technische Universit\"at Darmstadt, Germany and
%Professor at the Departments of Computer Science and of Mathematics.
Johannes Buchmann holds the Chair for Theoretical Computer Science (Cryptography
and Computer Algebra) at the department of Computer Science of the Technische
Universit\"at Darmstadt TUD).  He is also a Professor at the department of Mathematics, and vice-president of the university.

\item[Alexander May,] \mbox{}\\
Co-author of chapter~\ref{Chapter_Crypto2020} (``\nameref{Chapter_Crypto2020}''), 
Assistant Professor at the department of Computer Science of
the Technische Universit\"at Darmstadt, Germany.

\item[Erik Dahmen,] \mbox{}\\
Co-author of chapter~\ref{Chapter_Crypto2020} (``\nameref{Chapter_Crypto2020}''), 
%Researcher at the Department of Computer Science of the TUD.
Researcher at the Chair for Theoretical Computer Science (Cryptography and
Computer Algebra), department of Computer Science, Technische Universit\"at
Darmstadt, Germany.

\item[Ulrich Vollmer,] \mbox{}\\
Co-author of chapter~\ref{Chapter_Crypto2020} (``\nameref{Chapter_Crypto2020}''), 
Researcher at the Chair for Theoretical Computer Science (Cryptography and
Computer Algebra), department of Computer Science, Technische Universit\"at
Darmstadt, Germany.

@@

\newpage
\subsection{For Kids}

### 2009-11-27 script/en/chapters/cryptomethods.tex
Script[de+en]: Many small changes, e.g. Sage samples collected in files, use of iload; described how to deal with Sage notebook using interact; substituted href by url.

All classical methods are of this type. Examples can be found within the 
program CrypTool, in chapter  \ref{Kapitel_PaperandPencil} 
(Paper and Pencil Encryption Methods) of this script or in \cite{cm:Nichols1996}. 
Now we want to consider more modern mechanisms.

@@

More details can be found at:
\vspace{-10pt}
\begin{itemize}
  \item[] \href{http://www.cryptosystem.net/aes}
               {\texttt{http://www.cryptosystem.net/aes}}\\
          \href{http://www.minrank.org/aes/}
               {\texttt{http://www.minrank.org/aes/}}
\end{itemize}

@@

Companies like RSA Security provide so-called cipher challenges in order to quantify the security offered by well-known symmetric ciphers as DES, Triple-DES or RC5\footnote{\href{http://www.rsasecurity.com/rsalabs/challenges/secretkey/index.html}{\tt http://www.rsasecurity.com/rsalabs/challenges/secretkey/index.html}\\
In May 2007 RSA announced that they will not confirm the correctness of the not yet solved RC5-72 challenge.}. They offer prizes for those who manage to decipher ciphertexts, encrypted with different algorithms and different key lengths, and to unveil the symmetric key (under controlled conditions). So theoretical estimates can be confirmed.

It is well-known, that the ``old'' standard algorithm DES with a fixed key length of 56 bit is no more secure: this was demonstrated already in January 1999 by the Electronic Frontier Foundation (EFF). With their specialized computer Deep Crack they cracked a DES encrypted message within less than a day\footnote{\href{http://www.rsasecurity.com/rsalabs/challenges/des3/index.html}{\tt http://www.rsasecurity.com/rsalabs/challenges/des3/index.html}}.

@@

The RC5-64 challenge has been solved by the distributed.net team after 5 years\footnote{\href{http://distributed.net/pressroom/news-20020926.html}{\tt http://distributed.net/pressroom/news-20020926.html}}.  In total 331,252 individuals co-operated over the internet to find the key\footnote{%
An overview of current distributed computing projects can be found here:\\
\href{http://distributedcomputing.info/}{\tt http://distributedcomputing.info/}
}. More than 15 trillion (GB) / 15 quintillion (US)  keys were checked, until they found the right key.

@@

Similar cipher challenges are there for asymmetric algorithms (please see chapter \ref{NoteFactorization}).

>>>>

All classical methods are of this type. Examples can be found within the 
program CrypTool, in chapter~\ref{Kapitel_PaperandPencil}
(``\nameref{Kapitel_PaperandPencil}'')
of this script or in \cite{cm:Nichols1996}. 
Now we want to consider more modern mechanisms.

@@

More details can be found at:
\vspace{-10pt}
\begin{itemize}
  \item[] \url{http://www.cryptosystem.net/aes} \\
          \url{http://www.minrank.org/aes/} 
\end{itemize}

@@

Companies like RSA Security provide so-called cipher challenges in order to quantify the security offered by well-known symmetric ciphers as DES, Triple-DES or RC5\footnote{%
 \url{http://www.rsasecurity.com/rsalabs/challenges/secretkey/index.html}\\
 In May 2007 RSA announced that they will not confirm the correctness of
 the not yet solved RC5-72 challenge.}.
They offer prizes for those who manage to decipher ciphertexts, encrypted with different algorithms and different key lengths, and to unveil the symmetric key (under controlled conditions). So theoretical estimates can be confirmed.

It is well-known, that the ``old'' standard algorithm DES with a fixed key length
of 56 bit is no more secure: This was demonstrated already in January 1999 by the
Electronic Frontier Foundation (EFF). With their specialized computer Deep Crack
they cracked a DES encrypted message within less than a day\footnote{%
 \url{http://www.rsasecurity.com/rsalabs/challenges/des3/index.html}
}.

@@

The RC5-64 challenge has been solved by the distributed.net team after 5
years\footnote{%
 \url{http://distributed.net/pressroom/news-20020926.html}
}.
In total 331,252 individuals co-operated over the internet to find the key\footnote{%
An overview of current distributed computing projects can be found here:\\
\url{http://distributedcomputing.info/}
}.
More than 15 trillion (GB) / 15 quintillion (US)  keys were checked, until they
found the right key.

@@

There are also cipher challenges for asymmetric algorithms (please
see chapter \ref{NoteFactorization})\footnote{%
  A wide spectrum of simple and complex crypto riddles are included
  in the ``Mystery Twister C3''~challenge.
  \index{Challenge}
  \index{Cipher challenge}
  \index{Crypto challenge}
  }
.

### 2009-11-27 script/en/chapters/ellipticcurves.tex
Script[de+en]: Many small changes, e.g. Sage samples collected in files, use of iload; described how to deal with Sage notebook using interact; substituted href by url.

% -----------------------------------------------------------------------------
\newpage
% \section{Implementing elliptic curves for educational purposes, pat\-ent aspects}
\section{Implementing elliptic curves for educational purposes}

@@

\noindent There is an exhaustive, interactive ECC tutorial by Maike Massierer,
called \hyperlink{ec:Web-Link:Sage_Massierer}{ECC notebook}.
This interactive introduction to Elliptic Curve Cryptography is built up as
a Sage notebook.

@@

The ECC notebook\index{Elliptic curves!ECC notebook}\index{Massierer, Maike} of
Massierer\footnote{%
If you log on at another public server like
\url{http://sage.mathematik.uni-siegen.de:8000/home/pub/45/} or
\url{http://www.sagenb.org/}, and if there the according worksheets are installed
as "Published Worksheets", you can proceed as follows:\\
- Start the Sage notebook in a browser.

  \noindent\hangindent=6pt\makebox[6pt][l]{-}%
  Click in the above right corner on ``Published'', then you will see the 8 worksheets.

  \noindent\hangindent=6pt\makebox[6pt][l]{-}%
  Sometimes it's necessary at the beginning to click in the left upper corner on
  \verb#Action -> Evaluate all#.

  \noindent\hangindent=6pt\makebox[6pt][l]{-}%
  Sometimes the window contains many (blue) error messages. This normally can be
  solved by clicking the ``\%hide'' string: Then you get the code behind the
  graphics. Simply generate the graphics again with Shift-Enter.

  \noindent\hangindent=6pt\makebox[6pt][l]{-}%
  This tutorial uses the fonts from jsMath. They should be installed into your
  browser for a better layout.
  See \url{http://www.math.union.edu/~dpvc/jsMath/}
  and \url{http://pubpages.unh.edu/~jsh3/jsMath/}.\\
  After installing these fonts you can see the jsMath symbol at the lower border
  of your browser. If you click this symbol you can find the download page for
  the TIFF fonts. We are sorry, that this installation has to be done at every
  PC where this notebook runs in a browser.

  \noindent\hangindent=6pt\makebox[6pt][l]{-}%
  If you want to change one of the worksheets you should generate
  your own local copy (under Linux in your home area) by clicking
  ``Edit a copy'' in the left upper corner.

}${}^,$\footnote{%
  Since 2008 this ECC notebook can be found at
    \url{http://sage.mathematik.uni-siegen.de:8000/home/pub/45/}.\\
  To logon in Siegen you have to allow port 8000 and Cookies.\\
}
consists of 8 parts (title page plus 7 chapters)
% "Titelseite" mit Inhaltsverzeichnis plus 7 Kapitel
and aims to let even beginners understand what elliptic curves are:

>>>>

% -----------------------------------------------------------------------------
\newpage
% \section{Implementing elliptic curves for educational purposes, pat\-ent aspects}
\section{Implementing elliptic curves for educational purposes}
\label{ec:Implementing-for-Education}

@@

\noindent Additionally there is an exhaustive, interactive
\hyperlink{ec:Web-Link:Sage_Massierer}{ECC tutorial} by Maike Massierer.
This interactive introduction to Elliptic Curve Cryptography is built up as
a Sage notebook.

@@

The ECC notebook\index{Elliptic curves!ECC notebook}\index{Massierer, Maike} of
Massierer\footnote{%
Instructions to use an interactive Sage noebook:\\

  \noindent\hangindent=6pt\makebox[6pt][l]{-}%
  Public Sage servers like \url{http://sage.mathematik.uni-siegen.de:8000} or
  \url{http://www.sagenb.org/} often offer running samples as "Published Worksheets",
  which you can run and download without login on. These worksheets are listed
  if you click on ``Published'' in the above right corner.

  \noindent\hangindent=6pt\makebox[6pt][l]{-}%
  Worksheets using the  \verb#interact# command currently need some additional
  todos for a user to work correctly: sign-in, make a copy and execute all
  commands again.\\
  This works as follows (described for the sagenb server and for the ECC tutorial):\\

  \noindent\hangindent=6pt\makebox[6pt][l]{-}%
  Sign-up for a Sage notebook account at \url{http://sagenb.org/register}
  and sign in at \url{http://sagenb.org/}.

  \noindent\hangindent=6pt\makebox[6pt][l]{-}%
  Open the worksheet \url{http://sagenb.org/home/pub/1126/}.
  This contains the table of contents of the interactive ECC notebook.
  From here you can navigate via a click to the different chapters of the document.

  \noindent\hangindent=6pt\makebox[6pt][l]{-}%
  In the very top left corner, click \verb#Edit a copy# in order to create
  your own copy of the worksheet.

  \noindent\hangindent=6pt\makebox[6pt][l]{-}%
  Sometimes it's necessary at the beginning to re-evaluate the worksheet.
  Click in the left upper corner on \verb#Action -> Evaluate all#.

  \noindent\hangindent=6pt\makebox[6pt][l]{-}%
  Some of the applications still do not always work after opening a worksheet.
  Instead of nice output, they show lots of (blue) error messages.
  This normally can be solved quickly by clicking the gray  ``\%hide'' string:
  Then you get the code behind the graphics. Simply generate the graphics
  again with Shift-Enter.\\
  Even after doing this, the graphics code does not always disappear. Instead,
  it sometimes turns gray. Should this happen, click on the gray text,
  then click somewhere outside of the text box. The code will then
  disappear and leave you with a nice layout of the worksheet.

  \noindent\hangindent=6pt\makebox[6pt][l]{-}%
  Some of the ECC tutorial's content uses a special math fonts that are
  not installed by default with most browsers. When you notice that
  formulas are not displayed correctly or even get an error message
  about missing fonts from your browser, you need to install the jsMath
  fonts for a better layout.\\
  See \url{http://www.math.union.edu/~dpvc/jsMath/}
  and \url{http://pubpages.unh.edu/~jsh3/jsMath/}.\\
  After installing these fonts you can see the jsMath symbol at the lower border
  of your browser. If you click this symbol you can find the download page for
  the TIFF fonts. This fonts installation has to be done at every PC.

  \noindent\hangindent=6pt\makebox[6pt][l]{-}%
  According to the Sage-support newsgroup there is work underway
  to create a system for using @interact completely outside of the
  Sage notebook (JS code within a static html pages).

}${}^,$\footnote{%
  Since 2008 this ECC notebook can be found at
    \url{http://sage.mathematik.uni-siegen.de:8000/home/pub/45/} (\verb!#45 to #52!).
  To logon in Siegen you have to allow port 8000 and Cookies.\\
  Since 2009 an updated version of this ECC notebook can be found at
    \url{http://sagenb.org/home/pub/1126/} (\verb!#1126 to #1133!).
}
consists of 8 parts (title page plus 7 chapters)
% "Titelseite" mit Inhaltsverzeichnis plus 7 Kapitel
and aims to let even beginners understand what elliptic curves are:
